Return of the Linguist:\\Toward scalable parallel corpus generation for machine translation

### 2016-06-25 doc/paper.tex
just one more hack
various new doc ideas for amta
adverbial preposition infrastructure
add indirect object MOST of the time, at randomizer level
various comments

% conditional compilation
\def\buildemnlp{0}
\def\buildamta{1}
\def\buildhcomp{2}
\def\buildtarget{\buildemnlp} % required to be a single character for \if

% emnlp2016.sty, included by header-emnlp.tex, does NOT tolerate ANY surrounding \else statements...
% an if/else if/... or ifcase statement would have been preferred...
% but \documentclass is declared in the header file, which results in some hidden error-checking:
    % unsupported \buildtarget will fail to build (``\usepackage before \documentclass'')
    % duplicate target values (e.g., emnlp=amta=0) will fail to build (``two \documentclass commands'')
\if \buildtarget \buildemnlp    
    \input{header-emnlp.tex}
    \newcommand\mycitep[1]{\cite{#1}}     % conference-specific ``glue code''
    \newcommand\mynewcite[1]{\newcite{#1}}
    \newcommand\amtaonly[1]{} % currently being used to trim down to 4 pages
    \newcommand\emnlponly[1]{#1}
    \newcommand\hcomponly[1]{}
\fi
\if \buildtarget \buildamta    
    \input{header-amta.tex}
    \newcommand\mycitep[1]{\citep{#1}} 
    \newcommand\mynewcite[1]{\cite{#1}}  
    \newcommand\amtaonly[1]{#1}
    \newcommand\emnlponly[1]{}
    \newcommand\hcomponly[1]{}
\fi
\if \buildtarget \buildhcomp    
    \input{header-aaai.tex}
    \newcommand\mycitep[1]{\cite{#1}} 
    \newcommand\mynewcite[1]{\citeauthor{#1}~\shortcite{#1}}  
    \newcommand\amtaonly[1]{}
    \newcommand\emnlponly[1]{}
    \newcommand\hcomponly[1]{#1}
\fi

@@

% TODO (AMTA padding?): it's rather like having an athlete with an Olympic-level modern training regimen, when it comes to diet/nutrition, they'll just eat whatever they can find on the streets.
% Long and rich tradition of decades, if not centuries of linguists, "chefs" - want to exploit this somehow

@@

@@

% TODO: a reference to ``blessing of dimensionality'' for the inverse problem? how it's relatively easy to generate an exponentially large number of sentences?
Other practical considerations shine a favorable light on parallel corpus generation.
As the ``inverse'' of translation, multilingual generation is also an easier task in some senses.
For example, lexical ambiguity can be largely alleviated,\footnote{
    Ideally the generator would ``know'' {\em a priori} whether it is writing a sentence about, say, (financial) banks or (river) banks.
} segmentation of languages like Chinese \mycitep{segmenter:2005,segmenter:2008} becomes a non-issue, and there is the potential for generating annotated corpora with zero annotation error.
Parallel corpora also provide a natural avenue to interface contemporary MT systems with the countless linguist-years of expertise that have been invested into rule-based MT systems. %, as discussed in Section \ref{sec:related}. % that's not really discussed there anymore, is it?
And unlike the various different MT methods, which are by nature competitive, corpora are inherently collaborative.

@@

The idea of using synthetic parallel corpora in machine translation is hardly new.
Previous attempts at creating synthetic parallel corpora have used rule-based MT (RBMT) systems \mycitep{hu:2007,dugast:2008,murakami:2009,rubino:2014}, but the input to the RBMT systems was again dictated by external corpora in the desire to produce immediate results.
A longer-term approach that may pay dividends (Section \ref{subsec:benefits}) would be to run RBMT systems on specially chosen sentences for which they are known to produce high-quality translations, or perhaps eventually directly encoding their rules into carefully crafted parallel corpora.

@@

% POTENTIAL REVIEWER QUESTION (assuming they even understand wtf I was talking about here): well, but you can add word cluster ids as factors/features... (http://www.statmt.org/moses/?n=Advanced.Models#ntoc6 - simply annotate words )
% my weak responses
    % supervised data almost always trumps unsupervised data, especially in NLP. there is no real substitute for REAL, in-domain data
    % also, you would still have to require the models to to learn from these noisy clusterings
    % actually, word+mkcls is still a LEXICALIZED translation factor; that's probably why it doesn't really help during the translation phase
        % class-LM (target sequence model over cluster-ids) seems like it might be more useful....BUT, this is for MONOLINGUAL target-side data, 
            % which generally doesn't need the help
            % which is a strange thing to do to begin with, taking clusterings that were trained on expensive bilingual corpora and using them on plentiful monolingual corpora
% a possibly stronger response: automatic clusters are also of lower quality than hand-crafted template variations, and there will likely be clustering errors
% side note: if mkcls clusters were so awesome, they'd be used as standard factors... so you'd have to think that they're not

@@

% looks like this is GOING to run over length. save some if not all of this for AMTA, which is going to run way UNDER length
%\subsection{Simplistic Linguistics}
We adopt an extremely simplistic view of language as nouns, verbs, and their modifications and elaborations.
\amtaonly{
Inspired by early work on transformations by \mynewcite{chomsky}, we focus on declarative indicative present active sentences in the hopes that they can be transformed into other forms in the long run. 
Similarly, although we enforce some semantic constraints, our system generally produces nonsensical sentences like Chomsky's celebrated ``{\em Colorless green ideas sleep furiously.}''
}The fact that verb phrases can also serve as nouns (gerundives) and modifiers (participles) appears to be a major source of syntactic complexity, and we made it a priority to implement participial modifiers.
To capture recursive structures like participles, we adopt a rudimentary tree representation that is something intermediate between constituency and dependency trees, with constituent-like modifiers that also maintain dependency-like links to their targets.
We plan on eventually open-sourcing both our generator code and its accompanying data.

@@

%\subsection{Task}

@@

@@

% TODO: mention that official results are on a slightly different Moses configuration? mehhhhh
\caption{ \label{tab:results} % oh, for reference numbering, label goes INSIDE caption
    (Left) Chinese-to-English BLEU scores for our baseline Moses configuration with phrase and reordering tables trained on a random 25\% subset of the IWSLT15 training corpus and interpolated (+) with tables for the remaining 75\% (\textasciitilde 150K) of the in-domain corpus, the MultiUN corpus (a \textasciitilde 1M line subset), and a synthetic corpus from our generator (\textasciitilde 5M lines).
    (Right) Same as the left, but starting with tables trained on the entire IWSLT15 corpus.
    Official published scores of the IWSLT15 baseline system are also given for comparison.    
    Qualitatively similar trends for NIST and TER scores were also observed (omitted for clarity).
%    The ``baseline'' row gives 
%    MultiUN is actually more data because it has more words per line.
% automatically cleaned version of the en-zh_zh language pair of OpenSubtitles 2016, roughly 1M lines.
% and a similar number of lines from MultiUN, which is actually more words. but it doesn't do as well.
} % caption

@@

@@

>>>>

% conditional compilation
\def\buildemnlp{0}
\def\buildamta{1}
\def\buildhcomp{2}
\def\buildtarget{\buildamta} % required to be a single character for \if

\title{Return of the Linguist:\\Toward scalable parallel corpus generation for machine translation}

% emnlp2016.sty, included by header-emnlp.tex, does NOT tolerate ANY surrounding \else statements...
% an if/else if/... or ifcase statement would have been preferred...
% but \documentclass is declared in the header file, which results in some hidden error-checking:
    % unsupported \buildtarget will fail to build (``\usepackage before \documentclass'')
    % duplicate target values (e.g., emnlp=amta=0) will fail to build (``two \documentclass commands'')
\if \buildtarget \buildemnlp    
    \input{header-emnlp.tex}
    \newcommand\mycitep[1]{\cite{#1}}     % conference-specific ``glue code''
    \newcommand\mynewcite[1]{\newcite{#1}}
    \newcommand\amtaonly[1]{} % currently being used to trim down to 4 pages
    \newcommand\emnlponly[1]{#1}
    \newcommand\hcomponly[1]{}
\fi
\if \buildtarget \buildamta    
    \input{header-amta.tex}
    \newcommand\mycitep[1]{\citep{#1}} 
    \newcommand\mynewcite[1]{\cite{#1}}  
    \newcommand\amtaonly[1]{#1}
    \newcommand\emnlponly[1]{}
    \newcommand\hcomponly[1]{}

    % needed by Multeval tables
    \usepackage{amstext}
    \usepackage{multirow}
\fi
\if \buildtarget \buildhcomp    
    \input{header-aaai.tex}
    \newcommand\mycitep[1]{\cite{#1}} 
    \newcommand\mynewcite[1]{\citeauthor{#1}~\shortcite{#1}}  
    \newcommand\amtaonly[1]{}
    \newcommand\emnlponly[1]{}
    \newcommand\hcomponly[1]{#1}
\fi

@@

% TODO (AMTA padding?): it's rather like having an athlete with an Olympic-level modern training regimen, when it comes to diet/nutrition, they'll just eat whatever they can find on the streets.
% Long and rich tradition of decades, if not centuries of linguists, "chefs" - want to exploit this somehow
% fueled by data

@@

\amtaonly{

\begin{table}[htb]
\begin{center}
%\begin{footnotesize}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\bf Metric & \bf System & \bf Avg & \bf $\overline{s}_{\text{sel}}$ & \bf $s_{\text{Test}}$ & \bf $p$-value \\
\hline
\multirow{2}{*}{BLEU $\uparrow$}
& baseline & 11.5 & 0.3 & - & - \\
& system 1 & 10.9 & 0.3 & - & 0.00 \\
\hline
\multirow{2}{*}{TER $\downarrow$}
& baseline & 76.0 & 0.8 & - & - \\
& system 1 & 82.8 & 1.0 & - & 0.00 \\
\hline
\end{tabular}
%\end{footnotesize}
\end{center}
\caption{\label{tab:scores} Metric scores for all systems generated using MultEval 0.5.1 \mycitep{multeval}: jBLEU V0.1.1 (an exact reimplementation of NIST's mteval-v13.pl without tokenization); Translation Error Rate (TER) V0.8.0. p-values are relative to baseline and indicate whether a difference of this magnitude (between the baseline and the system on that line) is likely to be generated again by some random process (a randomized optimizer). Metric scores are averages over multiple runs. $s_{sel}$ indicates the variance due to test set selection and has nothing to do with optimizer instability.}
\end{table}

} % amtaonly

@@

% TODO: corpus generation is (almost like?) a sub-problem
% TODO: a reference to ``blessing of dimensionality'' for the inverse problem? how it's relatively easy to generate an exponentially large number of sentences?
% TODO: tree-based methods become a lot easier too? I even construct a random stupid tree myself...
Other practical considerations shine a favorable light on parallel corpus generation.
As the ``inverse'' of translation, multilingual generation is also an easier task in some senses.
For example, lexical ambiguity can be largely alleviated,\footnote{
    Ideally the generator would ``know'' {\em a priori} whether it is writing a sentence about, say, (financial) banks or (river) banks.
} segmentation of languages like Chinese \mycitep{segmenter:2005,segmenter:2008} becomes a non-issue, and there is the potential for generating annotated corpora with zero annotation error.
Parallel corpora also provide a natural avenue to interface contemporary MT systems with the countless linguist-years of expertise that have been invested into rule-based MT systems. %, as discussed in Section \ref{sec:related}. % that's not really discussed there anymore, is it?
And unlike the various different MT methods, which are by nature competitive, corpora are inherently collaborative.

@@

The idea of using synthetic parallel corpora in machine translation is hardly new.
% TODO: it's not domain limitation that matters - it's the fact that they're not amplifying, just mapping sentences 1-1
Previous attempts at creating synthetic parallel corpora have used rule-based MT (RBMT) systems \mycitep{hu:2007,dugast:2008,murakami:2009,rubino:2014}, but the input to the RBMT systems was again dictated by external corpora in the desire to produce immediate results.
A longer-term approach that may pay dividends (Section \ref{subsec:benefits}) would be to run RBMT systems on specially chosen sentences for which they are known to produce high-quality translations, or perhaps eventually directly encoding their rules into carefully crafted parallel corpora.

@@

% POTENTIAL REVIEWER QUESTION (assuming they even understand wtf I was talking about here): well, but you can add word cluster ids as factors/features... (http://www.statmt.org/moses/?n=Advanced.Models#ntoc6 - simply annotate words )
% my weak responses
    % supervised data almost always trumps unsupervised data, especially in NLP. there is no real substitute for REAL, in-domain data
    % also, you would still have to require the models to to learn from these noisy clusterings
    % actually, word+mkcls is still a LEXICALIZED translation factor; that's probably why it doesn't really help during the translation phase
        % class-LM (target sequence model over cluster-ids) seems like it might be more useful....BUT, this is for MONOLINGUAL target-side data, 
            % which generally doesn't need the help
            % which is a strange thing to do to begin with, taking clusterings that were trained on expensive bilingual corpora and using them on plentiful monolingual corpora
% a possibly stronger response: automatic clusters are also of lower quality than hand-crafted template variations, and there will likely be clustering errors
% side note: if mkcls clusters were so awesome, they'd be used as standard factors... so you'd have to think that they're not
% TODO: propose ``next-generation mkcls''? meh that's so ridiculously vague...

@@

% looks like this is GOING to run over length. save some if not all of this for AMTA, which is going to run way UNDER length
% TODO: maybe these nonsense sentences shed some light as to what is NOT a sensible, ``natural'' sentence...
%\subsection{Simplistic Linguistics}
We adopt an extremely simplistic view of language as nouns, verbs, and their modifications and elaborations.
\amtaonly{
Inspired by early work on transformations by \mynewcite{chomsky}, we focus on declarative indicative present active sentences in the hopes that they can be transformed into other forms in the long run. 
Similarly, although we enforce some semantic constraints, our system generally produces nonsensical sentences like Chomsky's celebrated ``{\em Colorless green ideas sleep furiously.}''
}The fact that verb phrases can also serve as nouns (gerundives) and modifiers (participles) appears to be a major source of syntactic complexity, and we made it a priority to implement participial modifiers.
To capture recursive structures like participles, we adopt a rudimentary tree representation that is something intermediate between constituency and dependency trees, with constituent-like modifiers that also maintain dependency-like links to their targets.
We plan on eventually open-sourcing both our generator code and its accompanying data.

@@

% TODO: mention that LM was held constant - since monolingual (even in-domain?) corpora are typically available at large scale
%\subsection{Task}

@@

% TODO: propose using sub-sentence timing information? you already HAVE relatively fine-grained alignment information... but the corpus crawler just throws that away. Then again, you might get stuff like cross-fragment alignments...

@@

% TODO: mention that official results are on a slightly different Moses configuration? mehhhhh
\caption{ \label{tab:results} % oh, for reference numbering, label goes INSIDE caption
    (Left) Chinese-to-English BLEU scores for our baseline Moses configuration with phrase and reordering tables trained on a random 25\% subset of the IWSLT15 training corpus and interpolated (+) with tables for the remaining 75\% (\textasciitilde 150K) of the in-domain corpus, the MultiUN corpus (a \textasciitilde 1M line subset), and a synthetic corpus from our generator (\textasciitilde 5M lines).
    (Right) Same as the left, but starting from tables trained on the entire IWSLT15 corpus.
    Official published scores of the IWSLT15 baseline system are also given for comparison.    
    Qualitatively similar trends for NIST and TER scores were also observed (omitted for clarity).
%    The ``baseline'' row gives 
%    MultiUN is actually more data because it has more words per line.
% automatically cleaned version of the en-zh_zh language pair of OpenSubtitles 2016, roughly 1M lines.
% and a similar number of lines from MultiUN, which is actually more words. but it doesn't do as well.
} % caption

@@

% TODO: current SMT is still really expert-sourced to professional translators. crowdsourcing translations takes things to their logical extreme.

@@

% TODO: I suspect MT requirements are very similar to that of, say, AP Latin - where the best translations are the most LITERAL, not necessarily the most fluent. 
% That also means that you might have to create TWO parallel corpora per language pair? To translate idioms in the source as compositional in the target...
% this would be closer to training data, since human translators are lazy too.

% TODO: Tarzan English (see OneNote)
